---
title: "Project1V100"
author: "Josh Horowitz"
date: "2/22/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
data1 <- read.csv("citibike_jan_sept_v1.csv")

library(dplyr)
library("ggplot2")
```


#What is the most frequently used start and end stations? 
####*start station: Pershing Squre North; end station: Pershing Square North*
```{r}

t <- table(data1$start.station.name)
w <- as.data.frame(t)
w[which.max(w$Freq),] 


t2 <- table(data1$end.station.name)
w2 <- as.data.frame(t2)
w2[which.max(w2$Freq),] 

```

#What is the most asymmetric start and end station? 
##The most assymetric start station is the station that has the greatest difference between start station frequency and end station frequency. In the output, the most asymmetric start station is the station with the highest positive number for "Freq_Dif". Similarly, the most assymetric end station is the station that has the greatest difference between end station frequency and start station frequency. In the output, the most end asymmetric station is the station with the highest negative number for "Freq_Dif". 
####*The most asymmetric start station is "Penn Station Valet". It has 98 more uses as a start station than as an end station. The most asymmetric end station is "Greenwhich Ave & 8 Ave". It has 75 more uses as an end station than as a start station.*
```{r}

#finding most asymmetric stations 

w <- as.data.frame(w)
w2 <- as.data.frame(w2)


merged_w_w2 <- full_join(w,w2, by = "Var1")
merged_w_w2 <- as.data.frame(merged_w_w2)

#adds new column (Freq_dif)
merged_w_w2$Freq_dif <- merged_w_w2$Freq.x - merged_w_w2$Freq.y 
# negative numbers in Freq_diff means more returns then starts 

#returns max diff ---most asymmetric station in favor of starts (most starts - ends)
most_start_asymm <- merged_w_w2[which.max(merged_w_w2$Freq_dif),] 
most_start_asymm

#returns min diff --most asymmetric station in favor of ends (most ends - starts)
most_end_asymm <- merged_w_w2[which.min(merged_w_w2$Freq_dif),]
most_end_asymm

```


#Below, there is a barplot showing the top 10 most assmyetric start stations. 
```{r, fig.height=10}

top_n <- function(dataFrame) {
  return(nrow(dataFrame) * .017)
}

#data frame of top 10 asymmetric start stations 
most_start_asymm_top10 <- merged_w_w2[order(merged_w_w2$Freq_dif, decreasing=TRUE)[1:top_n(merged_w_w2)],]

par(oma = c(10,0,0,0))

#plot of top 10 asymmetric start stations
barplot(most_start_asymm_top10$Freq_dif, names.arg = most_start_asymm_top10$Var1, col = c("slategray2", "slategray3", "slategray4"), las = 3, main = "Most Asymmetric Stations (Start Station > End Station)",  ylab = "# of times favoring start station usage over end station usage")

```

#Below, there is a barplot showing the top 10 most assymetric end stations. As the y-axis gets more negative, asymmetry increases. 
```{r, fig.height=10}
#data frame of top 10 asymmetric end stations 
most_end_asymm_top10 <- merged_w_w2[order(merged_w_w2$Freq_dif, decreasing=F)[1:top_n(merged_w_w2)],]

#plot of top 10 asymmetric end stations

par(oma = c(10,0,0,0))

barplot(most_end_asymm_top10$Freq_dif, names.arg = most_end_asymm_top10$Var1, col = c("slategray2", "slategray3", "slategray4"), las = 2, main = "Most Asymmetric Stations (End Station > Start Station)", ylab = "# of times favoring end station usage over start station usage")


```

#Does station frequency change by time of day? - Morning / Afternon / Evening
####*As the data below shows, frequency of start and end stations does change with time of day. While "Pershing Square North" was the the frequently used start and end station overall, "Penn Station Valet" was the most frequently used start station in the morning and "Pershing Square North" was the most frequently used end station in the morning. Also, "E 17 St. & Broadway" is the most frequenrly used start station in the afternoon, and "West St & Chambers St" is most frequently used end station in the afternoon. Lastly, "Pershing Square North" is the most used start and end station in the evening.*
```{r}

#morning data set 
data_morning <- subset(data1, data1$timeCategory == "Morning")
data_morning <- as.data.frame(data_morning)

#"Penn Station Valet" is most popular start station in the morning 
t_morning <- table(data_morning$start.station.name)
w_morning <- as.data.frame(t_morning)
w_morning[which.max(w_morning$Freq),]

#"Pershing Sqaure North is most popular end station in the morning"
t2_morning <- table(data_morning$end.station.name)
w2_morning <- as.data.frame(t2_morning)
w2_morning[which.max(w2_morning$Freq),] 


#afternoon data set
data_afternoon <- subset(data1, data1$timeCategory == "Afternoon")
data_afternoon <- as.data.frame(data_afternoon)

#"E 17 St. & Broadway" is most popular start station in the afternoon
t_afternoon <- table(data_afternoon$start.station.name)
w_afternoon <- as.data.frame(t_afternoon)
w_afternoon[which.max(w_afternoon$Freq),]

#"West St & Chambers St" is most popilar end station in the afternoon
t2_afternoon <- table(data_afternoon$end.station.name)
w2_afternoon <- as.data.frame(t2_afternoon)
w2_afternoon[which.max(w2_afternoon$Freq),] 


#evening data set 
data_evening <- subset(data1, data1$timeCategory == "Evening")
data_evening <- as.data.frame(data_evening)

#"Pershing Sqaure North" is most popular start station in the evening
t_evening <- table(data_evening$start.station.name)
w_evening <- as.data.frame(t_evening)
w_evening[which.max(w_evening$Freq),]

#"Pershing Square North" is most popular end station in the evening
t2_evening <- table(data_evening$end.station.name)
w2_evening <- as.data.frame(t2_evening)
w2_evening[which.max(w2_evening$Freq),] 

```

#Below, there is a barplot of the most frequently used start station for each time of the day. As the plot shows, the time of day that has the overall most frequently used start station is the evening. 
```{r}

#start station frequency barplot by time of day
start_station_freqs_time_day <- rbind(w_morning[which.max(w_morning$Freq),], w_afternoon[which.max(w_afternoon$Freq),], w_evening[which.max(w_evening$Freq),])
start_station_freqs_time_day$TimeDay <- c("Morning", "Afternoon", "Evening")
barplot(start_station_freqs_time_day$Freq, names.arg = start_station_freqs_time_day$TimeDay, col = c("slategray2", "slategray3", "slategray4"), las = 1, main = "Most Frequently Used Start Station by Time of Day", xlab = "Time of Day", ylab = "Frequency")

```

#Below, there is a barplot of the most frequently used end station for each time of the day. As the plot shows, the time of day that has the overall most frequently used end station is the evening.
```{r}

#end station frequency barplot by time of day
end_station_freqs_time_day <- rbind(w2_morning[which.max(w2_morning$Freq),], w2_afternoon[which.max(w2_afternoon$Freq),], w2_evening[which.max(w2_evening$Freq),])
end_station_freqs_time_day$TimeDay <- c("Morning", "Afternoon", "Evening")
barplot(end_station_freqs_time_day$Freq, names.arg = end_station_freqs_time_day$TimeDay, col = c("slategray2", "slategray3", "slategray4"), las = 1, main = "Most Frequently Used End Station by Time of Day", xlab = "Time of Day", ylab = "Frequency")

```


#Does station asymmetry change by time of day? - Morning / Afternoon / Evening
####*As the data below shows, start and end station asymmetry changes by time of day."Penn Station Valet" is the overall most asymmetric start station and "Greenwhich Ave & 8 Ave" is the overall most asymmetric end station. "Penn Station Valet" is the most asymmetric start station in the morning, while "W 52 St & 5 Ave" is the most asymmetric end station in the morning. "Penn Station Valet" is the most asymmetric start station in the afteroon, while "Centre St & Chambers St" is the most asymmetric end station in the afternoon. "W 52 St & 5 Ave" is the most asymmetric start station in the evening, while "	E 14 St & Avenue B" is the most asymmetric end station in the evening.*
```{r}

#morning
w_morning <- as.data.frame(w_morning)
w2_morning <- as.data.frame(w2_morning)

merged_w_w2_morning <- full_join(w_morning,w2_morning, by = "Var1")
merged_w_w2_morning <- as.data.frame(merged_w_w2_morning)

#adds new column (Freq_dif)
merged_w_w2_morning$Freq_dif <- merged_w_w2_morning$Freq.x - merged_w_w2_morning$Freq.y 
# negative numbers in Freq_diff means more returns then starts 

#returns max diff ---most asymmetric station in favor of starts (most starts - ends)
most_start_asymm_morning <- merged_w_w2_morning[which.max(merged_w_w2_morning$Freq_dif),] 
most_start_asymm_morning

#returns min diff --most asymmetric station in favor of ends (most ends - starts)
most_end_asymm_morning <- merged_w_w2_morning[which.min(merged_w_w2_morning$Freq_dif),]
most_end_asymm_morning

#afternoon
w_afternoon <- as.data.frame(w_afternoon)
w2_afternoon <- as.data.frame(w2_afternoon)

merged_w_w2_afternoon <- full_join(w_morning,w2_afternoon, by = "Var1")
merged_w_w2_afternoon <- as.data.frame(merged_w_w2_afternoon)

#adds new column (Freq_dif)
merged_w_w2_afternoon$Freq_dif <- merged_w_w2_afternoon$Freq.x - merged_w_w2_afternoon$Freq.y 
# negative numbers in Freq_diff means more returns then starts 

#returns max diff ---most asymmetric station in favor of starts (most starts - ends)
most_start_asymm_afternoon <- merged_w_w2_afternoon[which.max(merged_w_w2_afternoon$Freq_dif),] 
most_start_asymm_afternoon

#returns min diff --most asymmetric station in favor of ends (most ends - starts)
most_end_asymm_afternoon <- merged_w_w2_afternoon[which.min(merged_w_w2_afternoon$Freq_dif),]
most_end_asymm_afternoon

#evening
w_evening <- as.data.frame(w_evening)
w2_evening <- as.data.frame(w2_evening)

merged_w_w2_evening <- full_join(w_evening,w2_evening, by = "Var1")
merged_w_w2_evening <- as.data.frame(merged_w_w2_evening)

#adds new column (Freq_dif)
merged_w_w2_evening$Freq_dif <- merged_w_w2_evening$Freq.x - merged_w_w2_evening$Freq.y 
# negative numbers in Freq_diff means more returns then starts 

#returns max diff ---most asymmetric station in favor of starts (most starts - ends)
most_start_asymm_evening <- merged_w_w2_evening[which.max(merged_w_w2_evening$Freq_dif),] 
most_start_asymm_evening

#returns min diff --most asymmetric station in favor of ends (most ends - starts)
most_end_asymm_evening <- merged_w_w2_evening[which.min(merged_w_w2_evening$Freq_dif),]
most_end_asymm_evening

```

#Below, there is a barplot of the highest asymmetric start station for each time of day. As the barplot shows, the time of day with the highest start station asymmetry is the morning, and the the time of day with the lowest start startion asymmetry is the evening. 
```{r}

#barplot of most assymetric start station by time of day 
most_start_asymm_time_day <- rbind(most_start_asymm_morning, most_start_asymm_afternoon, most_start_asymm_evening)

most_start_asymm_time_day$time_of_day <- c("Morning", "Afternoon", "Evening")


barplot(most_start_asymm_time_day$Freq_dif, names.arg = most_start_asymm_time_day$time_of_day, col = c("slategray2", "slategray3", "slategray4"), las = 1, main = "Most Asymmetric Start Station by Time of Day", xlab = "Time of Day", ylab = "Asymmetry Count")

```

#Below, there is a barplot of the highest asymmetric end station for each time of day. As the barplot suggests, end station asymmetry is the highest in the morning, second highest in the evening, and least in the afternoon. 
```{r}

#barplot of most assymetreic end station by time of day
most_end_asymm_time_day <- rbind(most_end_asymm_morning, most_end_asymm_afternoon, most_end_asymm_evening)

most_end_asymm_time_day$time_of_day <- c("Morning", "Afternoon", "Evening")


barplot(most_end_asymm_time_day$Freq_dif, names.arg = most_end_asymm_time_day$time_of_day, col = c("slategray2", "slategray3", "slategray4"), las = 1, main = "Most Asymmetric End Station by Time of Day", xlab = "Time of Day", ylab = "Asymmetry Count")

```

#Does start and end station frequency change by day of week?
####*As the data below shows, start and end station frequency change with some days of the week. "Pershing Square North" is the most frequently used start and end station overall. The most frequently used start station on Monday is "West St & Chambers St", while the most frequently used end station on Monday is "Pershing Square North". The most frequently used start and end station on Tuesday, Wednesday, Thursday,and Friday is "Pershing Square North". On Saturday, "West St & Chambers St" is the most frequently used start and end station. Lastly,  "W 21 St & 6 Ave is the most frequently used start station on Sunday and "Centre St & Chambers St" is the most frequently used end station on Sunday.*
```{r}

#Monday data set 
data_monday <- subset(data1, data1$weekday == "Monday")
data_monday <- as.data.frame(data_monday)

#"West St & Chambers St" is most popular start station on Monday
t_monday <- table(data_monday$start.station.name)
w_monday <- as.data.frame(t_monday)
w_monday[which.max(w_monday$Freq),]

#"Pershing Square North"" is most popular end station on Monday
t2_monday <- table(data_monday$end.station.name)
w2_monday <- as.data.frame(t2_monday)
w2_monday[which.max(w2_monday$Freq),]

#Tuesday data set 
data_tuesday <- subset(data1, data1$weekday == "Tuesday")
data_tuesday <- as.data.frame(data_tuesday)

#"Pershing Square North" is most popular start station on Tuesday
t_tuesday <- table(data_tuesday$start.station.name)
w_tuesday <- as.data.frame(t_tuesday)
w_tuesday[which.max(w_tuesday$Freq),]


#"Pershing Sqaure North" is most popular end station on Tuesday
t2_tuesday <- table(data_tuesday$end.station.name)
w2_tuesday <- as.data.frame(t2_tuesday)
w2_tuesday[which.max(w2_tuesday$Freq),]


#Wedneday data set 
data_wed <- subset(data1, data1$weekday == "Wednesday")
data_wed <- as.data.frame(data_wed)

#"Pershing Square North" is most popular start station on Wednesday
t_wed <- table(data_wed$start.station.name)
w_wed <- as.data.frame(t_wed)
w_wed[which.max(w_wed$Freq),]

#"Pershing Sqaure North" is most popular end station on Wednesday
t2_wed <- table(data_wed$end.station.name)
w2_wed <- as.data.frame(t2_wed)
w2_wed[which.max(w2_wed$Freq),]

#Thurday data set 
data_thurs <- subset(data1, data1$weekday == "Thursday")
data_thurs <- as.data.frame(data_thurs)

#"Pershing Square North" is most popular start station on Thursday
t_thurs <- table(data_thurs$start.station.name)
w_thurs <- as.data.frame(t_thurs)
w_thurs[which.max(w_thurs$Freq),]

#"Pershing Square North" is most popular end station on Thursday
t2_thurs <- table(data_thurs$end.station.name)
w2_thurs <- as.data.frame(t2_thurs)
w2_thurs[which.max(w2_thurs$Freq),] 

#Friday data set 
data_fri <- subset(data1, data1$weekday == "Friday")
data_fri <- as.data.frame(data_fri)

#"Pershing Square North" is most popular start station on Friday
t_fri <- table(data_fri$start.station.name)
w_fri <- as.data.frame(t_fri)
w_fri[which.max(w_fri$Freq),]

#"Pershing Square North" is most popular end station on Friday
t2_fri <- table(data_fri$end.station.name)
w2_fri <- as.data.frame(t2_fri)
w2_fri[which.max(w2_fri$Freq),]

#Saturday data set 
data_sat <- subset(data1, data1$weekday == "Saturday")
data_sat <- as.data.frame(data_sat)

#"West St & Chambers St" is most popular start station on Saturday
t_sat <- table(data_sat$start.station.name)
w_sat <- as.data.frame(t_sat)
w_sat[which.max(w_sat$Freq),]

#"West St & Chambers St" is most popular end station on Saturday
t2_sat <- table(data_sat$end.station.name)
w2_sat <- as.data.frame(t2_sat)
w2_sat[which.max(w2_sat$Freq),]

#Sunday data set 
data_sun <- subset(data1, data1$weekday == "Sunday")
data_sun <- as.data.frame(data_sun)

#"W 21 St & 6 Ave" is most popular start station on Sunday
t_sun <- table(data_sun$start.station.name)
w_sun <- as.data.frame(t_sun)
w_sun[which.max(w_sun$Freq),]

#"Centre St & Chambers St" is most popular end station on Sunday
t2_sun <- table(data_sun$end.station.name)
w2_sun <- as.data.frame(t2_sun)
w2_sun[which.max(w2_sun$Freq),] 

```

#Below, there are barplots of the most frequently used start station and end station for each day of the week. As the plots show, the day of the week that has the most frequently used start station is Thursday, and the day of the week that has the most frequently used end station is Tuesday. 
```{r}
#barplot of most frequent used start station for each day of week
start_station_freq_day_week <- rbind(w_monday[which.max(w_monday$Freq),], w_tuesday[which.max(w_tuesday$Freq),], w_wed[which.max(w_wed$Freq),], w_thurs[which.max(w_thurs$Freq),], w_fri[which.max(w_fri$Freq),], w_sat[which.max(w_sat$Freq),], w_sun[which.max(w_sun$Freq),])

start_station_freq_day_week$Day_of_Week <- c("Monday", "Tuesday", "Wed", "Thursday", "Friday", "Saturday", "Sunday")

barplot(start_station_freq_day_week$Freq, names.arg = start_station_freq_day_week$Day_of_Week, col = c("slategray2", "slategray3", "slategray4", "steelblue2", "steelblue3", "steelblue4", "lightblue4"), las = 1, main = "Frequency of Most Frequently Used Start Station for each Day of Week", xlab = "Day of Week", ylab = "Frequency")


#barplot of most frequent used end station for each day of week
end_station_freq_day_week <- rbind(w2_monday[which.max(w2_monday$Freq),], w2_tuesday[which.max(w2_tuesday$Freq),], w2_wed[which.max(w2_wed$Freq),], w2_thurs[which.max(w2_thurs$Freq),], w2_fri[which.max(w2_fri$Freq),], w2_sat[which.max(w2_sat$Freq),], w2_sun[which.max(w2_sun$Freq),])

end_station_freq_day_week$Day_of_Week <- c("Monday", "Tuesday", "Wed", "Thursday", "Friday", "Saturday", "Sunday")

barplot(end_station_freq_day_week$Freq, names.arg = end_station_freq_day_week$Day_of_Week, col = c("slategray2", "slategray3", "slategray4", "steelblue2", "steelblue3", "steelblue4", "lightblue4"), las = 1, main = "Frequency of Most Frequently Used End Station for each Day of Week", xlab = "Day of Week", ylab = "Frequency")
```

#Does Start and End Station Asymmetry change with the day of the week?
####*As the data shows, start and end station asymmetry change with the day of the week.The most asymmetric start station on Monday is "Grand Army Plaza & Central Park S", with 21 more uses as a start station than as an end station. The most asymmetric end station on Monday is "W 41 St & 8 Ave", with 18 more uses as an end station than as a start station.The most asymmetric start station on Tuesday is "Penn Station Valet", with 31 more uses as a start station than as an end station. The most asymmetric end station on Tuesday is "W 39 St & 9 Ave", with 24 more uses as an end station than as a start station.The most asymmetric start station on Wednesday is "W 42 St & Dyer Ave", with 26 more uses as a start station than as an end station. The most asymmetric end station on Wednesday is "Cleveland Pl & Spring St", with 24 more uses as an end station than as a start station.The most asymmetric start station on Thursday is "Pershing Square North", with 26 more uses as a start station than as an end station. The most asymmetric end station on Thursday is "Broadway & W 29 St", with 22 more uses as an end station than as a start station.The most asymmetric start station on Friday is "Broadway & W 29 St", with 18 more uses as a start station than as an end station. The most asymmetric end station on Friday is "W 41 St & 8 Ave", with 41 more uses as an end station than as a start station.The most asymmetric start station on Saturday is "Grand Army Plaza & Central Park S", with 17 more uses as a start station than as an end station. The most asymmetric end station on Saturday is "W 21 St & 6 Ave", with 19 more uses as an end station than as a start station.The most asymmetric start station on Sunday is "W 20 St & 11 Ave", with 16 more uses as a start station than as an end station. The most asymmetric end station on Sunday is "Centre St & Chambers St", with 21 more uses as an end station than as a start station.*
```{r}

#monday
w_monday <- as.data.frame(w_monday)
w2_monday <- as.data.frame(w2_monday)


merged_w_w2_monday <- full_join(w_monday,w2_monday, by = "Var1")
merged_w_w2_monday <- as.data.frame(merged_w_w2_monday)

#adds new column (Freq_dif)
merged_w_w2_monday$Freq_dif <- merged_w_w2_monday$Freq.x - merged_w_w2_monday$Freq.y 
# negative numbers in Freq_diff means more returns then starts 

#returns max diff ---most asymmetric station in favor of starts (most starts - ends)
most_start_asymm_monday <- merged_w_w2_monday[which.max(merged_w_w2_monday$Freq_dif),] 
most_start_asymm_monday

#returns min diff --most asymmetric station in favor of ends (most ends - starts)
most_end_asymm_monday <- merged_w_w2_monday[which.min(merged_w_w2_monday$Freq_dif),]
most_end_asymm_monday

#tuesday
w_tuesday <- as.data.frame(w_tuesday)
w2_tuesday<- as.data.frame(w2_tuesday)


merged_w_w2_tuesday <- full_join(w_tuesday,w2_tuesday, by = "Var1")
merged_w_w2_tuesday <- as.data.frame(merged_w_w2_tuesday)

#adds new column (Freq_dif)
merged_w_w2_tuesday$Freq_dif <- merged_w_w2_tuesday$Freq.x - merged_w_w2_tuesday$Freq.y 
# negative numbers in Freq_diff means more returns then starts 

#returns max diff ---most asymmetric station in favor of starts (most starts - ends)
most_start_asymm_tuesday <- merged_w_w2_tuesday[which.max(merged_w_w2_tuesday$Freq_dif),] 
most_start_asymm_tuesday

#returns min diff --most asymmetric station in favor of ends (most ends - starts)
most_end_asymm_tuesday <- merged_w_w2_tuesday[which.min(merged_w_w2_tuesday$Freq_dif),]
most_end_asymm_tuesday

#wednesday
w_wed <- as.data.frame(w_wed)
w2_wed <- as.data.frame(w2_wed)


merged_w_w2_wed <- full_join(w_wed,w2_wed, by = "Var1")
merged_w_w2_wed <- as.data.frame(merged_w_w2_wed)

#adds new column (Freq_dif)
merged_w_w2_wed$Freq_dif <- merged_w_w2_wed$Freq.x - merged_w_w2_wed$Freq.y 
# negative numbers in Freq_diff means more returns then starts 

#returns max diff ---most asymmetric station in favor of starts (most starts - ends)
most_start_asymm_wed <- merged_w_w2_wed[which.max(merged_w_w2_wed$Freq_dif),] 
most_start_asymm_wed

#returns min diff --most asymmetric station in favor of ends (most ends - starts)
most_end_asymm_wed <- merged_w_w2_wed[which.min(merged_w_w2_wed$Freq_dif),]
most_end_asymm_wed 

#thursday
w_thurs <- as.data.frame(w_thurs)
w2_thurs <- as.data.frame(w2_thurs)


merged_w_w2_thurs <- full_join(w_thurs,w2_thurs, by = "Var1")
merged_w_w2_thurs <- as.data.frame(merged_w_w2_thurs)

#adds new column (Freq_dif)
merged_w_w2_thurs$Freq_dif <- merged_w_w2_thurs$Freq.x - merged_w_w2_thurs$Freq.y 
# negative numbers in Freq_diff means more returns then starts 

#returns max diff ---most asymmetric station in favor of starts (most starts - ends)
most_start_asymm_thurs <- merged_w_w2_thurs[which.max(merged_w_w2_thurs$Freq_dif),] 
most_start_asymm_thurs

#returns min diff --most asymmetric station in favor of ends (most ends - starts)
most_end_asymm_thurs <- merged_w_w2_thurs[which.min(merged_w_w2_thurs$Freq_dif),]
most_end_asymm_thurs

#friday
w_fri <- as.data.frame(w_fri)
w2_fri <- as.data.frame(w2_fri)


merged_w_w2_fri <- full_join(w_fri,w2_fri, by = "Var1")
merged_w_w2_fri <- as.data.frame(merged_w_w2_fri)

#adds new column (Freq_dif)
merged_w_w2_fri$Freq_dif <- merged_w_w2_fri$Freq.x - merged_w_w2_fri$Freq.y 
# negative numbers in Freq_diff means more returns then starts 

#returns max diff ---most asymmetric station in favor of starts (most starts - ends)
most_start_asymm_fri <- merged_w_w2_fri[which.max(merged_w_w2_fri$Freq_dif),] 
most_start_asymm_fri

#returns min diff --most asymmetric station in favor of ends (most ends - starts)
most_end_asymm_fri <- merged_w_w2_fri[which.min(merged_w_w2_fri$Freq_dif),]
most_end_asymm_fri

#sat
w_sat <- as.data.frame(w_sat)
w2_sat <- as.data.frame(w2_sat)


merged_w_w2_sat <- full_join(w_sat,w2_sat, by = "Var1")
merged_w_w2_sat <- as.data.frame(merged_w_w2_sat)

#adds new column (Freq_dif)
merged_w_w2_sat$Freq_dif <- merged_w_w2_sat$Freq.x - merged_w_w2_sat$Freq.y 
# negative numbers in Freq_diff means more returns then starts 


#returns max diff ---most asymmetric station in favor of starts (most starts - ends)
most_start_asymm_sat <- merged_w_w2_sat[which.max(merged_w_w2_sat$Freq_dif),] 
most_start_asymm_sat

#returns min diff --most asymmetric station in favor of ends (most ends - starts)
most_end_asymm_sat <- merged_w_w2_sat[which.min(merged_w_w2_sat$Freq_dif),]
most_end_asymm_sat

#sun
w_sun <- as.data.frame(w_sun)
w2_sun <- as.data.frame(w2_sun)


merged_w_w2_sun <- full_join(w_sun,w2_sun, by = "Var1")
merged_w_w2_sun <- as.data.frame(merged_w_w2_sun)

#adds new column (Freq_dif)
merged_w_w2_sun$Freq_dif <- merged_w_w2_sun$Freq.x - merged_w_w2_sun$Freq.y 
# negative numbers in Freq_diff means more returns then starts 

#returns max diff ---most asymmetric station in favor of starts (most starts - ends)
most_start_asymm_sun <- merged_w_w2_sun[which.max(merged_w_w2_sun$Freq_dif),] 
most_start_asymm_sun

#returns min diff --most asymmetric station in favor of ends (most ends - starts)
most_end_asymm_sun <- merged_w_w2_sun[which.min(merged_w_w2_sun$Freq_dif),]
most_end_asymm_sun

```

#Below, there are barplots of stations with the largest start and end asymmetry for each day of the week. As the first the plot shows, the day with largest amount of start station asymmetry is Tuesday. As the second plot shows, the day with the most end station asymmety is Friday. 
```{r}

#barplot of most asymmetric start station by day of week 
most_start_asymm_day_week <- rbind(most_start_asymm_monday, most_start_asymm_tuesday, most_start_asymm_wed, most_start_asymm_thurs, most_start_asymm_fri, most_start_asymm_sat, most_start_asymm_sun)

most_start_asymm_day_week$Day_of_Week <- c("Monday", "Tuesday", "Wed", "Thursday", "Friday", "Saturday", "Sunday")

barplot(most_start_asymm_day_week$Freq_dif, names.arg = most_start_asymm_day_week$Day_of_Week, col = c("slategray2", "slategray3", "slategray4", "steelblue2", "steelblue3", "steelblue4", "lightblue4"), las = 1, main = "Most Asymmetric Start Station by Day of Week", xlab = "Day of Week", ylab = "Asymmetry Count") 


#barplot of most asymmetric end station by day of week 
most_end_asymm_day_week <- rbind(most_end_asymm_monday, most_end_asymm_tuesday, most_end_asymm_wed, most_end_asymm_thurs, most_end_asymm_fri, most_end_asymm_sat, most_end_asymm_sun)

most_end_asymm_day_week$Day_of_Week <- c("Monday", "Tuesday", "Wed", "Thursday", "Friday", "Saturday", "Sunday")

barplot(most_end_asymm_day_week$Freq_dif, names.arg = most_end_asymm_day_week$Day_of_Week, col = c("slategray2", "slategray3", "slategray4", "steelblue2", "steelblue3", "steelblue4", "lightblue4"), las = 1, main = "Most Asymmetric End Station by Day of Week", xlab = "Day of Week", ylab = "Asymmetry Count")


```

